---
title: "Brennan_Test"
author: "Brennan Thompson"
date: "2024-08-03"
output: html_document
---

### MLR 2: Can we predict the amount needed for financial independence based off of the lifestyle one leads?

#### Importing Required Libraries

```{r}
library(tidyverse)
library(MASS)
library(broom)
library("ggpubr")
```
#### BRENNAN DO THIS: You need to clean the data, turn it info a df that you then use to create a model after getting rid of the periods and eliminating the columns. You use this to calculate residuals. 


#### Reading the Data

```{r}
work_dir <- dirname(getwd())
filepath <- paste(work_dir, 
                  "/data/reddit_finance - vish - renamed_year_cols.csv", 
                  sep = "") 
fin_df <- read.csv(filepath)
```

#### Pre-Processing & Filters

```{r}
# only US Sample
fin_df <- filter(fin_df, country == "United States")

# Response Variable -> fin_indy_inst_amount
# Didn't make sense to replace it with 0's so we replaced it with the mean. People might not have understood the question and so just left it blank as FIRE is all one thing, so they wouldn't have considered financial independence as independent from retirement
# Calculate means, ignoring NA values
mean_fin_indy_num <- mean(fin_df$fin_indy_num, na.rm = TRUE)
mean_whn_fin_indy_num <- mean(fin_df$whn_fin_indy_num, na.rm = TRUE)

# Replace NA's with the mean value and sum the columns
fin_df <- fin_df %>% 
  mutate(fin_indy_inst_amount = 
           coalesce(fin_indy_num, mean_fin_indy_num) + 
           coalesce(whn_fin_indy_num, mean_whn_fin_indy_num))

# Filter the amount to be less than 10000000
fin_df <- filter(fin_df, fin_indy_inst_amount < 10000000)

# Square the output
fin_df <- fin_df %>% mutate(fin_indy_inst_amount = (1 + fin_indy_inst_amount)**(1/4))

# Filtering to avoid division by zero
fin_df <- filter(fin_df, is.finite(fin_indy_inst_amount))
# Plotting histogram
fin_df %>% ggplot(aes(x = fin_indy_inst_amount)) + geom_histogram()

# QQ plot
ggqqplot(fin_df$fin_indy_inst_amount)
```
```{r}
# There is a very faint U-shape, which isn't promising
# Get the number of values in the fin_indy_inst_amount column
num_rows <- length(fin_df$fin_indy_inst_amount)
```

```{r}

# Print the number of values
print(num_rows)
```
#### Major Transformations

```{r}
### BASIC TRANSFORMATIONS
model_df <- fin_df %>% mutate(
  tgt_sf_wthdrw_rt = coalesce(as.numeric(tgt_sf_wthdrw_rt),3),
  tgt_sf_wthdrw_rt = ifelse(tgt_sf_wthdrw_rt < 15, tgt_sf_wthdrw_rt, NA),
  tgt_sf_wthdrw_rt = coalesce(tgt_sf_wthdrw_rt, mean(tgt_sf_wthdrw_rt)),
  
  retire_exp = coalesce(as.numeric(retire_exp),0),
  retire_exp = ifelse(retire_exp==0, mean(retire_exp), retire_exp),
  
  home_value = coalesce(home_value,0),
  
  brokerage_accts_tax = coalesce(brokerage_accts_tax,0),
  brokerage_accts_tax = ifelse(brokerage_accts_tax!=0,
                               brokerage_accts_tax, 
                               mean(brokerage_accts_tax)),
  
  retirement_accts_tax = coalesce(retirement_accts_tax, 0),
  retirement_accts_tax = ifelse(retirement_accts_tax==0, 
                                mean(retirement_accts_tax), 
                                retirement_accts_tax),
  
  cash = coalesce(cash, 0),
  cash = ifelse(cash==0 | cash>100000, mean(cash), cash),
  
  invst_accts = coalesce(invst_accts, 0),
  invst_accts = ifelse(invst_accts==0 | invst_accts>100000, 
                       mean(invst_accts), 
                       invst_accts),
  
  spec_crypto = coalesce(spec_crypto, 0),
  
  invst_prop_bus_own = coalesce(invst_prop_bus_own,0),
  
  student_loans = coalesce(student_loans,0),
  
  other_val = coalesce(other_val,0),
  
  auto_loan = coalesce(auto_loan,0),
  
  credit_personal_loan = coalesce(credit_personal_loan,0),
  
  medical_debt = coalesce(medical_debt,0),
  
  Y2020_gross_inc = coalesce(Y2020_gross_inc,0),
  Y2020_housing_exp = coalesce(Y2020_housing_exp,0),
  Y2020_utilities_exp = coalesce(Y2020_utilities_exp,0),
  Y2020_transp_exp = coalesce(Y2020_transp_exp,0),
  Y2020_necessities_exp = coalesce(Y2020_necessities_exp,0),
  Y2020_lux_exp = coalesce(Y2020_lux_exp,0),
  Y2020_child_exp = coalesce(Y2020_child_exp,0),
  Y2020_debt_repay = coalesce(Y2020_debt_repay,0),
  Y2020_invst_save = coalesce(Y2020_invst_save,0),
  Y2020_charity = coalesce(Y2020_charity,0),
  Y2020_healthcare_exp = coalesce(Y2020_healthcare_exp,0),
  Y2020_taxes = coalesce(Y2020_taxes,0),
  Y2020_edu_exp = coalesce(Y2020_edu_exp,0),
  Y2020_other_exp = coalesce(Y2020_other_exp,0),
  retire_invst_num = coalesce(retire_invst_num,0)
  

  
)
```

```{r}
### A few more linear transformations
model_df <- model_df %>% mutate(
  invst_prop_bus_own = log (1+invst_prop_bus_own),
  Y2020_lux_exp = sqrt(1 + Y2020_lux_exp),
  Y2020_gross_inc = sqrt(1 + Y2020_gross_inc),
  retire_invst_num = log(1+ retire_invst_num)
  )

model_df = dplyr::select(model_df, c(retire_invst_num,
                              
    # children,
    
    fin_indy_inst_amount,
    tgt_sf_wthdrw_rt,
    retire_exp,
    home_value,
    brokerage_accts_tax,
    retirement_accts_tax,
    cash,
    invst_accts,
    spec_crypto,
    invst_prop_bus_own,
    other_val,
    student_loans,
    auto_loan,
    credit_personal_loan,
    medical_debt,

    Y2020_gross_inc,
    Y2020_housing_exp,
    Y2020_utilities_exp,
    Y2020_transp_exp,
    Y2020_necessities_exp,
    Y2020_lux_exp,
    Y2020_child_exp,
    Y2020_debt_repay,
    Y2020_invst_save,
    Y2020_charity,
    Y2020_healthcare_exp,
    Y2020_taxes,
    Y2020_edu_exp,
    Y2020_other_exp,

    
    

)
    
  
)
```

```{r}
model_df2 <- gather(model_df, key = "predictor", value = "value", -fin_indy_inst_amount)

```

```{r}
model_df2 %>%
  ggplot(
    aes(y = fin_indy_inst_amount, x = value)
  ) + 
  geom_jitter() + 
  facet_wrap(predictor ~ ., scale = "free") + 
  geom_smooth(method = "lm")

```

#### OLS MODEL

```{r}
base_model <- lm(fin_indy_inst_amount~., model_df)
```

```{r}
bmsumm = summary(base_model)
```

```{r}
bmsumm
```


```{r}
bmvif <- car::vif(base_model)
```


```{r}
bmvif
```

#### OUTLIER REMOVAL

```{r}
outliers = filter(bm_diag, abs(.std.resid)>=3)
model_df_2 = filter(bm_diag, abs(.std.resid)<3)
model_df_2 = dplyr::select(model_df_2, -starts_with("."))
```

```{r}
model_df_2
```

#### Re-Run Base OLS Model

```{r}
base_model2 <- lm(fin_indy_inst_amount~., model_df_2)
```

```{r}
bmsumm2 = summary(base_model2)
```

```{r}
bmsumm2
```

```{r}
bmvif2 <- car::vif(base_model2)
```

```{r}
bmvif2
```

#### AKAKI INFORMATION CRITERIA

```{r}
aic_model <- MASS::stepAIC(base_model2, direction = "both", trace = F) 
summary(aic_model)
```

#### Removing High p-value Features

```{r}
low_lm <- lm(fin_indy_inst_amount ~ . -Y2020_healthcare_exp,data = model_df_2)
```

```{r}
aic_model <- MASS::stepAIC(low_lm, direction = "both", trace = F) 
summary(aic_model)
```

```{r}
low_lm <- lm(fin_indy_inst_amount ~ . -Y2020_healthcare_exp - home_value, data = model_df_2)
```

```{r}
aic_model <- MASS::stepAIC(low_lm, direction = "both", trace = F) 
summary(aic_model)
```

```{r}
low_lm <- lm(fin_indy_inst_amount ~ . -Y2020_healthcare_exp - home_value - invst_accts , data = model_df_2)
```

```{r}
aic_model <- MASS::stepAIC(low_lm, direction = "both", trace = F) 
summary(aic_model)
```

```{r}
bmvif3 <- car::vif(aic_model)
```

```{r}
bmvif3
```

```{r}
model_df = dplyr::select(model_df, c(
'fin_indy_inst_amount',                              
"retire_invst_num", 
  "tgt_sf_wthdrw_rt", 
  "retire_exp", 
  "brokerage_accts_tax", 
  "retirement_accts_tax", 
  "cash",
  "spec_crypto", 
  "student_loans", 
  "Y2020_gross_inc", 
  "Y2020_housing_exp", 
  "Y2020_debt_repay", 
  "Y2020_invst_save"
)
    
  
)

```

```{r}
model_df2 <- gather(model_df, key = "predictor", value = "value", -fin_indy_inst_amount)
```

```{r}
model_df2 %>%
  ggplot(
    aes(y = fin_indy_inst_amount, x = value)
  ) + 
  geom_jitter() + 
  facet_wrap(predictor ~ ., scale = "free") + 
  geom_smooth(method = "lm")

```

```{r}
# Need to check for and remove outliers, maybe even graph the dang things
# OUTLIER REMOVAL NEXT LINE
```


```{r}
diagnostics2 <- aic_model %>%
  augment(data = na.omit(model_df_2)) 
diagnostics2 %>% ggplot(aes(x = .fitted,y = .resid)) + geom_jitter()
```

