---
title: "MLR1"
author: "VG_qtf7du"
date: "2024-08-05"
output: html_document
---

### MLR 2: Can we predict the Amount of Investment required at retirement based on the Lifestyle one leads?

#### Importing Required Libraries

```{r}
library(tidyverse)
library(MASS)
library(broom)
library("ggpubr")
library(caret)
```

#### Reading the Data

```{r}
work_dir <- dirname(getwd())
filepath <- paste(work_dir, 
                  "/data/reddit_finance - vish - renamed_year_cols.csv", 
                  sep = "") 
fin_df_raw <- read.csv(filepath)
```

#### Pre-Processing & Filters

```{r}
# only US Sample
fin_df <- filter(fin_df_raw, country == "United States")

# Response Variable -> retire_invst_num
fin_df <- filter(fin_df, 
                 coalesce(retire_invst_num,0)>1000 & 
                 coalesce(retire_invst_num,0)<10000000)

fin_df <- fin_df %>% mutate(retire_invst_num_raw = (retire_invst_num))
fin_df <- fin_df %>% mutate(retire_invst_num = (retire_invst_num)**(0.5))

# Cleaning the Fin Indy Amount for both pops
fin_df <- fin_df %>% 
  mutate(fin_indy_inst_amount = 
           coalesce(fin_indy_num, 0) + 
           coalesce(whn_fin_indy_num, 0),
         
         fin_indy_inst_amount = ifelse(fin_indy_inst_amount==0, 
                                       mean(fin_indy_inst_amount),
                                       fin_indy_inst_amount))

fin_df <- filter(fin_df, fin_indy_inst_amount<5000000)


fin_df %>% ggplot(aes(x = retire_invst_num_raw, y= fin_indy_inst_amount)) + geom_jitter()


ggqqplot(fin_df$retire_invst_num) + ggtitle("Normality of Response Variable - retire_invst_num")
```

```{r}
fin_df <- mutate(fin_df, own.the.house = as.factor(ifelse(housing == "Own", "Yes","No")))
summary(fin_df$homer)
# gather(dplyr::select(fin_df,c(retire_invst_num,fin_indy_inst_amount,homer)), key = "predictor", value = "value", -homer)  %>% 
#   ggplot(aes(x = value, color=homer, fill=homer))  + facet_wrap(predictor ~ ., scale = "free") + geom_density()

# dplyr::filter(fin_df, homer == 0) %>% ggplot(aes(x = retire_invst_num, y= fin_indy_inst_amount,color = homer)) + geom_jitter()
fin_df %>% ggplot(aes(x = retire_invst_num_raw, y= fin_indy_inst_amount,color = own.the.house)) + geom_jitter() + geom_smooth(method = "glm") + ggtitle("Relationship between FI v RE & classification on owning of home") +ylab("FI Number") + xlab("RE Number")

```


#### Major Transformations

```{r}

### BASIC TRANSFORMATIONS
fin_df <- fin_df %>% mutate(
  tgt_sf_wthdrw_rt = coalesce(as.numeric(tgt_sf_wthdrw_rt),7.5),
  tgt_sf_wthdrw_rt = ifelse(tgt_sf_wthdrw_rt < 15, tgt_sf_wthdrw_rt, NA),
  tgt_sf_wthdrw_rt = coalesce(tgt_sf_wthdrw_rt, mean(tgt_sf_wthdrw_rt)),
  tgt_sf_wthdrw_rt = coalesce(as.numeric(tgt_sf_wthdrw_rt),7.5),
  
  retire_exp = coalesce(as.numeric(retire_exp),0),
  retire_exp = ifelse(retire_exp==0, mean(retire_exp), retire_exp),
  
  tgt_sf_wthdrw_val = retire_exp/tgt_sf_wthdrw_rt,
  
  home_value = coalesce(as.numeric(home_value),0),
  home_value = ifelse(home_value==0, mean(home_value)/2, home_value),
  
  brokerage_accts_tax = coalesce(brokerage_accts_tax,0),
  brokerage_accts_tax = ifelse(brokerage_accts_tax!=0,
                               brokerage_accts_tax, 
                               mean(brokerage_accts_tax)),
  
  retirement_accts_tax = coalesce(retirement_accts_tax, 0),
  retirement_accts_tax = ifelse(retirement_accts_tax==0, 
                                mean(retirement_accts_tax), 
                                retirement_accts_tax),
  
  cash = coalesce(cash, 0),
  cash = ifelse(cash==0 | cash>100000, mean(cash), cash),
  
  invst_accts = coalesce(invst_accts, 0),
  invst_accts = ifelse(invst_accts==0 | invst_accts>100000, 
                       mean(invst_accts), 
                       invst_accts),
  
  spec_crypto = coalesce(spec_crypto, 0),
  invst_prop_bus_own = coalesce(invst_prop_bus_own,0),
  student_loans = coalesce(student_loans,0),
  other_val = coalesce(other_val,0),
  auto_loan = coalesce(auto_loan,0),
  credit_personal_loan = coalesce(credit_personal_loan,0),
  medical_debt = coalesce(medical_debt,0),
  
  Y2020_gross_inc = coalesce(Y2020_gross_inc,0),
  Y2020_housing_exp = coalesce(Y2020_housing_exp,0),
  Y2020_utilities_exp = coalesce(Y2020_utilities_exp,0),
  Y2020_transp_exp = coalesce(Y2020_transp_exp,0),
  Y2020_necessities_exp = coalesce(Y2020_necessities_exp,0),
  Y2020_lux_exp = coalesce(Y2020_lux_exp,0),
  Y2020_child_exp = coalesce(Y2020_child_exp,0),
  Y2020_debt_repay = coalesce(Y2020_debt_repay,0),
  Y2020_invst_save = coalesce(Y2020_invst_save,0),
  Y2020_charity = coalesce(Y2020_charity,0),
  Y2020_healthcare_exp = coalesce(Y2020_healthcare_exp,0),
  Y2020_taxes = coalesce(Y2020_taxes,0),
  Y2020_edu_exp = coalesce(Y2020_edu_exp,0),
  Y2020_other_exp = coalesce(Y2020_other_exp,0),
  other_debt = coalesce(other_debt,0),
  mortgage = coalesce(mortgage, 0)

  
)



fin_df <- fin_df %>% mutate(
  fin_indy_inst_amount = (fin_indy_inst_amount)**(1/2),
  tgt_sf_wthdrw_val =tgt_sf_wthdrw_val**0.5,
  retire_exp = retire_exp**0.5,
  home_value = sqrt(home_value),
  invst_accts = invst_accts**0.5,
  retirement_accts_tax = sqrt(retirement_accts_tax),
  invst_prop_bus_own = sqrt(invst_prop_bus_own),
  Y2020_lux_exp = sqrt(Y2020_lux_exp),
  Y2020_gross_inc = log(1 + Y2020_gross_inc),
  spec_crypto = log(1 + spec_crypto),
  brokerage_accts_tax = sqrt(brokerage_accts_tax),
  auto_loan = sqrt(auto_loan),
  cash = cash**0.5,
  credit_personal_loan = credit_personal_loan,
  # Y2020_utilities_exp = (Y2020_utilities_exp)**(1/3)
  
  )

model_df = dplyr::select(fin_df, c(retire_invst_num,
                              
    fin_indy_inst_amount,
    # tgt_sf_wthdrw_val,
    tgt_sf_wthdrw_rt,
    retire_exp,
    home_value,
    brokerage_accts_tax,
    retirement_accts_tax,
    cash,
    invst_accts,
    spec_crypto,
    invst_prop_bus_own,
    other_val,
    student_loans,
    auto_loan,
    credit_personal_loan,
    medical_debt,
    other_debt,
    mortgage,
    fin_indy_pct,
    


    Y2020_gross_inc,
    Y2020_housing_exp,
    Y2020_utilities_exp,
    Y2020_transp_exp,
    Y2020_necessities_exp,
    Y2020_lux_exp,
    Y2020_child_exp,
    Y2020_debt_repay,
    Y2020_invst_save,
    Y2020_charity,
    Y2020_healthcare_exp,
    Y2020_taxes,
    Y2020_edu_exp,
    Y2020_other_exp,
    
    
    children,
    edu,
    stp_whn_fin_indy
)
    
  
)
# dplyr::select(model_df,names(fin_df)[sapply(fin_df, is.numeric)])
# model_df2 <- gather(dplyr::select(model_df,names(fin_df)[sapply(fin_df, is.numeric)]), key = "predictor", value = "value", -retire_invst_num)

# 
# model_df2 %>%
#   ggplot(
#     aes(y = retire_invst_num, x = value)
#   ) + 
#   geom_jitter() + 
#   facet_wrap(predictor ~ ., scale = "free") + 
#   geom_smooth(method = "lm")
# 
# 
# blank <- apply(is.na(model_df), 1, any)
# blanks = model_df[blank, ]
# 
model_df <- na.omit(model_df)
# cor1 = cor(model_df)
# cor1
```

```{r}
nums <- data.frame(names(fin_df)[sapply(fin_df, is.numeric)])
```

#### OLS MODEL

```{r}
base_model <- lm(retire_invst_num~., model_df)
bmsumm = summary(base_model)
bmsumm


bmvif <- car::vif(base_model)
bmvif
```

#### AKAKI INFORMATION CRITERIA

```{r}
aic_model <- MASS::stepAIC(base_model, direction = "both", trace = F) 
summary(aic_model)
```

#### OUTLIER REMOVAL

```{r}
bm_diag <- aic_model %>%   augment(data = model_df)  
bm_diag    
outliers = filter(bm_diag, abs(.std.resid)>=3) 

model_df_2 = filter(bm_diag, abs(.std.resid)<3) 
# model_df_2 = filter(model_df_2, abs(.hat)<3) 
model_df_2 = dplyr::select(model_df_2, -starts_with(".")) 
model_df_2
```

#### Re-Run Base AKAKI Model

```{r}
base_model2 <- MASS::stepAIC(lm(retire_invst_num~., model_df_2) , direction = "both", trace = F) 
bmsumm2 = summary(base_model2) 
bmsumm2   
bmvif2 <- car::vif(base_model2) 
# bmvif2
```

#### 

#### Removing High p-value Features

```{r}
low_lm <- lm(retire_invst_num ~ .
              -Y2020_lux_exp      -Y2020_child_exp                                                                                                      # -Y2020_necessities_exp -Y2020_invst_save
             # -fin_indy_pct -home_value -tgt_sf_wthdrw_val
             ,data = model_df_2)
summary(low_lm)
aic_model <- MASS::stepAIC(low_lm, direction = "both", trace = F) 
aic_summary <- summary(aic_model)
data.frame(car::vif(aic_model))
```

```{r}
aic_summary
```

#### Regularization Attempt

```{r}
library(glmnet)
X = model.matrix(retire_invst_num~0+., data = model_df_2)
y = model_df_2$retire_invst_num

rmodel <- glmnet(x=X, y=y, alpha = 1)
plot(rmodel, label = T, xvar = "lambda")
```

```{r}
kfcvglmnet <- cv.glmnet(x=X, y=y, alpha=1, nfolds=3)
plot(kfcvglmnet, label = T, xvar = "lambda")
kfcvglmnet$lambda.min
```




#### Final Selected Model

```{r}
final_model_df = na.omit(dplyr::select(model_df_2, retire_invst_num,
                               
# fin_indy_inst_amount ,
# tgt_sf_wthdrw_rt ,
# retire_exp ,
# home_value ,
# brokerage_accts_tax ,
# retirement_accts_tax ,
# invst_prop_bus_own ,
# other_val ,
# student_loans ,
# credit_personal_loan ,
# medical_debt ,
# other_debt ,
# mortgage ,
# fin_indy_pct ,
# Y2020_gross_inc ,
# Y2020_housing_exp ,
# Y2020_utilities_exp ,
# Y2020_transp_exp ,
# Y2020_necessities_exp ,
# Y2020_lux_exp ,
# Y2020_debt_repay ,
# Y2020_healthcare_exp ,
# Y2020_taxes ,
# Y2020_edu_exp ,
# Y2020_other_exp ,



# fin_indy_inst_amount ,
# tgt_sf_wthdrw_rt ,
# retire_exp ,
# brokerage_accts_tax ,
# retirement_accts_tax ,
# cash ,
# auto_loan ,
# medical_debt ,
# other_debt ,
# mortgage ,
# Y2020_housing_exp ,
# Y2020_utilities_exp ,
# Y2020_lux_exp ,
# Y2020_healthcare_exp ,
# Y2020_taxes ,
# Y2020_edu_exp ,
# children,
# stp_whn_fin_indy,


# fin_indy_inst_amount ,
# tgt_sf_wthdrw_val ,
# tgt_sf_wthdrw_rt ,
# home_value ,
# brokerage_accts_tax ,
# cash ,
# invst_prop_bus_own ,
# other_debt ,
# Y2020_gross_inc ,
# Y2020_transp_exp ,
# Y2020_child_exp ,
# Y2020_taxes ,
# Y2020_edu_exp ,
# children,
# stp_whn_fin_indy





fin_indy_inst_amount ,
tgt_sf_wthdrw_rt ,
retire_exp ,
home_value ,
brokerage_accts_tax ,
cash ,
invst_prop_bus_own ,
other_debt ,
Y2020_gross_inc ,
Y2020_transp_exp ,
Y2020_invst_save ,
Y2020_taxes ,
Y2020_edu_exp ,
children,
stp_whn_fin_indy

))

final_model = lm(retire_invst_num ~ ., data = final_model_df)
summary(final_model)
diagnostics <- augment(final_model, data = na.omit(final_model_df))
diagnostics %>% ggplot(aes(x = .fitted,y = .resid)) + geom_jitter()
```
```{r}
summary(final_model)
```

#### Residual Analysis

```{r}
diagnostics <- augment(final_model, data = na.omit(final_model_df))
ggqqplot(diagnostics$.resid)
```

```{r}
data.frame(car::vif(final_model))
```


```{r}
library(ggcorrplot)

cor(dplyr::select(final_model_df,-c(children, stp_whn_fin_indy)))
ggcorrplot(cor(dplyr::select(final_model_df,-c(children, stp_whn_fin_indy))),
           type = "upper", lab = T
           ) + ggtitle("Correlation between major numeric predictors in MLR")
```

```{r}

model_df2 <- gather(dplyr::select(final_model_df,-c(children,stp_whn_fin_indy)), key = "predictor", value = "value", -retire_invst_num)


model_df2 %>%
  ggplot(
    aes(y = retire_invst_num, x = value)
  ) +
  geom_jitter() +
  facet_wrap(predictor ~ ., scale = "free") +
  geom_smooth(method = "lm")
```

```{r}
means <- sapply(final_model_df, mean, na.rm = TRUE)
 
# Brennan Example
user_inputs <- list(
  fin_indy_inst_amount = sqrt(3100000),
  student_loans = sqrt(25000),
  tgt_sf_wthdrw_rt = 3.00,
  spec_crypto = 80000,
  stp_whn_fin_indy = "Yes",
  children = "Do not have children, but intend to"
)

new_data <- data.frame(matrix(ncol = length(names(means)), nrow = 1))
colnames(new_data) <- names(means)

for (var in names(user_inputs)) {
  new_data[[var]] <- user_inputs[[var]]
}

for (var in setdiff(names(means), names(user_inputs))) {
  new_data[[var]] <- means[[var]]
}

prediction <- predict(final_model, newdata = new_data, interval = 'prediction')
print((prediction)**2)
```
#### CROSS VALIDATION 

```{r}
control <- trainControl(method = "repeatedcv", number = 7, repeats = 10)
model2 <- train(retire_invst_num~., method = "lm", trControl = control, data = final_model_df)
summary(model2)
model2$results$RMSE
```


