---
title: "Data Exploration and Research Question"
author: "Afnan Alabdulwahab"
date: "2024-07-31"
output: html_document
---

```{r}
#install.packages("DataExplorer")
# Load the DataExplorer package
# Install skimr and summarytools
#install.packages("skimr")
#install.packages("summarytools")

library(DataExplorer)
# Load the packages
library(skimr)
library(tidyverse)
```

```{r}
# Load the dataset
reddit_data <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/reddit_finance.csv")
```

```{r}
# Inspect unique values in the country column
unique(reddit_data$country)
```

```{r}
# Filter the dataset to include only respondents from the United States
us_reddit_data <- reddit_data %>% filter(country == "United States")

# View the first few rows of the filtered dataset
head(us_reddit_data)
```

```{r}
# Descriptive statistics for numeric variables
us_reddit_data %>%
  select_if(is.numeric) %>%
  summary()
```


```{r}
# Inspect unique values in the age column
unique(us_reddit_data$age)
```

```{r}
# Create a bar plot for age distribution
ggplot(filtered_us_reddit_data, aes(x = age)) +
  geom_bar(fill = "blue", color = "black") +
  labs(title = "Age Distribution of Respondents in the United States", x = "Age Range", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Create a bar plot for gender distribution
ggplot(us_reddit_data, aes(x = gender)) +
  geom_bar(fill = "purple", color = "black", na.rm = TRUE) +
  labs(title = "Distribution of Gender", x = "Gender", y = "Count")
```

```{r}
# Create a bar plot for education distribution
ggplot(us_reddit_data, aes(x = edu)) +
  geom_bar(fill = "orange", color = "black", na.rm = TRUE) +
  labs(title = "Distribution of Education Level", x = "Education Level", y = "Count")
```


```{r}
# Create a data profiling report
create_report(us_reddit_data)
```
```{r}
# Using skimr
skim(us_reddit_data)
```


I think this question might work for categorical predictions since age (age group) and education level are categorical
## Research Question 1: 
**Research Question:** How do age, education level, and number of incomes influence the Financial Independence (FI) number among American respondents?

**Response Variable:**
- `fin_indy_num`: The amount at which respondents consider themselves financially independent.

**Predictors:**
1. **Age**: Convert age ranges to numeric values (as midpoint values).
2. **Education Level**: Encode education levels as numeric or factor variables.
3. **Number of Incomes**: The number of individuals contributing to the household income (`num_incomes`).

```{r}
# Load the dataset
reddit_data <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/reddit_finance.csv")

# Filter the dataset to include only respondents from the United States
us_reddit_data <- reddit_data %>% filter(country == "United States")
```


```{r}
# Filter out rows with missing values in the relevant columns
cleaned_data <- us_reddit_data %>%
  filter(age != "N/A" & age != "" & !is.na(fin_indy_num) & !is.na(num_incomes) & !is.na(edu)) %>%
  mutate(
    age_factor = factor(age),                # Convert age to factor
    edu_factor = factor(edu, levels = unique(edu))  # Convert education to factor
  )
cleaned_data
```


```{r}
# Select relevant columns
model_data <- cleaned_data %>% dplyr::select(fin_indy_num, age_factor, num_incomes, edu_factor)
```

```{r}
# View summary of the data
summary(model_data)
```

```{r}
# Visualize relationships between predictors and response variable
ggplot(model_data, aes(x = age_factor, y = fin_indy_num)) + 
  geom_boxplot() + 
  labs(title = "Financial Independence Number vs Age", x = "Age Range", y = "FI Number")

ggplot(model_data, aes(x = num_incomes, y = fin_indy_num)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Financial Independence Number vs Number of Incomes", x = "Number of Incomes", y = "FI Number")

ggplot(model_data, aes(x = edu_factor, y = fin_indy_num)) + 
  geom_boxplot() + 
  labs(title = "Financial Independence Number vs Education Level", x = "Education Level", y = "FI Number")
```

